COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region

### 2020-03-25 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

```{r}
pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
```
A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
#Spatial model  
We estimate a BYM model fixing prior distributions for iid and besag components as follows:  
```{r}
prior.iid = c(1,0.01)
prior.besag = c(1,0.001)
initial.iid = 4
initial.besag = 3
```
We create new variables that are required by INLA procedure.  
```{r}
##########################indexing, for 
#### number of days
dat_csv$t<-as.numeric(dat_csv$data)
days<-max(dat_csv$t)
dat_csv$t3<-dat_csv$t2<-dat_csv$t
#### province
dat_csv$ID2<-dat_csv$ID
#### province
dat_csv$IDt<-1:dim(dat_csv)[1]
########################## formula BYM
formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_1)
##incidence rate ratio for each province
IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
nc.province$IRR_mean<-IRR_mean
nc.province$DEN_PROV<-droplevels(nc.province$DEN_PROV)

spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios", sp.layout=
         list("sp.text", coordinates(nc.province), nc.province$DEN_PROV, cex=0.7, col="gray"))

```
The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.

@@

##################predictions 3 days forward of new CODID 19 cases

```{r}
Forecast=3
dat_csv2<-dat_csv 
dat_csv2$t<-rep(1:days+days,nprov)
dat_csv2$totale_casi<-NA
dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
dat_csv_n<-rbind(dat_csv,dat_csv2)
dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
#introducting hubei starting on 22/01, quarantine started on 25/01
# applying a smoothing , data has errors
hubei_new_cases<-filter(diff(hubei), rep(1, 5))

plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
abline(v=2,col=1,lty=2)
abline(v=4,col=2,lty=2)
abline(v=10,col=3,lty=3)
legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
text(10,1500,"Hubei quarantine")
text(12,1350,"Lodi home lock",col=2)
text(20,1200,"Lombardy home lock",col=3)
#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
#other provinces on 11/03 lagged started their restrictions
hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
dat_csv_n$hubei<-rep(hubei_others,12)
dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
#overdispersion
dat_csv_n$over<-1:dim(dat_csv_n)[1]
```

@@

```{r}
## include pop e hubei
formula.int_n<-  totale_casi ~ 1+log(hubei+1)+log(pop)+ f(ID,model="bym",graph=file.adj) +
f(t,model="rw2") + f(t2,model="iid") +
f(t3,model="iid", group=ID2, control.group=list(model="besag", graph=file.adj))

fit_n<-inla(formula.int_n, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
summary(fit_n)
```

```{r}
#### components
#overall trend
t_1<-fit_n$summary.random$t$mean 
#iid trend
t_2<-fit_n$summary.random$t2$mean 
 # spatial CAR componen
s_1<-fit_n$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
# spatio-temporal coefficients 
st<-fit_n$summary.random$t3$mean
## Trends for province
Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
trends<-t_1+t(matrix(st,nrow=nprov))
plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date")
matplot(t(matrix(st,nrow=nprov)),ylab="RW2 Time coefficient for Province",xlab="Date")
#extract number of cases for lodi
est_lodi<-fit_n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]

days.before<-Date_n[1:days]
days.ahead<-Date_n[(days+1):(days+Forecast)]
mu.lower<-est_lodi$mean-1.96*est_lodi$sd
mu.lower[mu.lower<0]<-0
mu.upper<-est_lodi$mean+1.96*est_lodi$sd
mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
step.ahead<-Forecast
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

>>>>

```{r}
pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
```
A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
#Spatial model  
We estimate a BYM model fixing prior distributions for iid and besag components as follows:  
```{r}
prior.iid = c(1,0.01)
prior.besag = c(1,0.001)
initial.iid = 4
initial.besag = 3
```
We create new variables that are required by INLA procedure.  
```{r}
##########################indexing, for 
#### number of days
dat_csv$t<-as.numeric(dat_csv$data)
days<-max(dat_csv$t)
dat_csv$t3<-dat_csv$t2<-dat_csv$t
#### province
dat_csv$ID2<-dat_csv$ID
#### province
dat_csv$IDt<-1:dim(dat_csv)[1]
########################## formula BYM
formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_1)
##incidence rate ratio for each province
IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
nc.province$IRR_mean<-IRR_mean
nc.province$DEN_PROV<-droplevels(nc.province$DEN_PROV)
spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios", sp.layout=
         list("sp.text", coordinates(nc.province), nc.province$DEN_PROV, cex=0.7, col="gray"))
```
The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.

@@

##################predictions 7 days forward of new CODID 19 cases

```{r}
Forecast=7
dat_csv2<-dat_csv 
dat_csv2$t<-rep(1:days+days,nprov)
dat_csv2$totale_casi<-NA
dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
dat_csv_n<-rbind(dat_csv,dat_csv2)
dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
#introducting hubei starting on 22/01, quarantine started on 25/01
# applying a smoothing , data has errors
hubei_new_cases<-filter(diff(hubei), rep(1, 5))

#plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
#lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
#abline(v=2,col=1,lty=2)
#abline(v=4,col=2,lty=2)
#abline(v=10,col=3,lty=3)
#legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
#text(10,1500,"Hubei quarantine")
#text(12,1350,"Lodi home lock",col=2)
#text(20,1200,"Lombardy home lock",col=3)
#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
#other provinces on 11/03 lagged started their restrictions
hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
dat_csv_n$hubei<-rep(hubei_others,12)
dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
#overdispersion
dat_csv_n$over<-1:dim(dat_csv_n)[1]
```

@@

```{r}
## include pop e hubei
formula.int_n<-  totale_casi ~ log(pop)+hubei+f(ID,model="bym",graph=file.adj) +
f(t,model="rw2",constr = FALSE) + f(t2,model="iid") +
f(over,model="iid")

link = rep(NA, dim(dat_csv_n)[1])
link[which(is.na(dat_csv_n$totale_casi))] = 1

fit_n<-inla(formula.int_n, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = link))
summary(fit_n)
```

```{r}
#### components
#overall trend
t_1<-fit_n$summary.random$t$mean 
#iid trend
t_2<-fit_n$summary.random$t2$mean 
 # spatial CAR componen
s_1<-fit_n$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
# spatio-temporal coefficients 
st<-fit_n$summary.random$over$mean
## Trends for province
Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
trends<-t_1+t(matrix(st,ncol=(days+Forecast)))
plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date",ylim=range(trends),type="l",lwd=2)
matlines(Date_n,trends,type="l")
abline(v=Date_n[days],lty=2)
```

#Forecast for Lodi province

```{r}
est_lodi<-fit_n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]

days.before<-Date_n[1:days]
days.ahead<-Date_n[(days+1):(days+Forecast)]
mu.lower<-est_lodi$`0.025quant`
mu.upper<-est_lodi$`0.975quant`
mu.med<-xts(est_lodi$`0.5quant`,order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
step.ahead<-Forecast
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

### 2020-03-25 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

```{r}
names(dat_csv[,-c(1:2,13)])

@@

```{r}
myDateTimeStr1 <- dat_csv$data
myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
days_dy<-as.Date(myPOSIXct1)
dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
```

>>>>

```{r}
names(dat_csv[,-c(1:2,17:19)])

@@

```{r}
myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
days_dy<-as.Date(myPOSIXct1)
dat_csv_dy<-xts(dat_csv[,-c(1:2,13:15)], order.by = days_dy, frequency = 7)
```

### 2020-03-25 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

```{r}
myDateTimeStr1 <- dat_csv$data
myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
days_dy<-as.Date(myPOSIXct1)
dat_csv_dy<-xts(dat_csv[,-c(1:7,16:17)], order.by = days_dy, frequency = 7)
```

>>>>

```{r}
myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
days_dy<-as.Date(myPOSIXct1)
dat_csv_dy<-xts(dat_csv[,-c(1:7,17:19)], order.by = days_dy, frequency = 7)
```
